---
title: "Homework I"
author: "Parker Lutz"
date: "June 8, 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## II. b)

####(a) Plot the observations.

```{r a, comment=FALSE, warning=FALSE}
clusterData <- cbind(c(1, 1, 0, 5, 6, 4), c(4, 3, 4, 1, 2, 0))
plot(clusterData[,1], clusterData[,2])
```

####(b) Randomly assign a cluster label to each observation.
####Report the clusterlabels for each observation.

```{r b, comment=FALSE, warning=FALSE}
set.seed(1)
labels <- sample(2, nrow(clusterData), replace = T)
labels
```


####(c) Compute the centroid for each cluster.

```{r c, comment=FALSE, warning=FALSE}
plot(clusterData[, 1], clusterData[, 2], col = (labels + 1), pch = 20, ceclusterData = 2)
center_1 <- c(mean(clusterData[labels == 1, 1]), mean(clusterData[labels == 1, 2]))
center_1
center_2 <- c(mean(clusterData[labels == 2, 1]), mean(clusterData[labels == 2, 2]))
center_2
plot(clusterData[,1], clusterData[,2], col=(labels + 1), pch = 20, ceclusterData = 2)
points(center_1[1], center_1[2], col = 2, pch = 4)
points(center_2[1], center_2[2], col = 3, pch = 4)
```

####(d) Assign each observation to the centroid to which it is closest, in
####terms of Euclidean distance. Report the cluster labels for each
####observation.

```{r d, comment=FALSE, warning=FALSE}
labels <- c(1, 1, 1, 2, 2, 2)
plot(clusterData[, 1], clusterData[, 2], col = (labels + 1), pch = 20, ceclusterData = 2)
points(center_1[1], center_1[2], col = 2, pch = 4)
points(center_2[1], center_2[2], col = 3, pch = 4)
```

####(e) Repeat (c) and (d) until the answers obtained stop changing.

```{r e, comment=FALSE, warning=FALSE}
center_1 <- c(mean(clusterData[labels == 1, 1]), mean(clusterData[labels == 1, 2]))
center_2 <- c(mean(clusterData[labels == 2, 1]), mean(clusterData[labels == 2, 2]))
plot(clusterData[,1], clusterData[,2], col=(labels + 1), pch = 20, ceclusterData = 2)
points(center_1[1], center_1[2], col = 2, pch = 4)
points(center_2[1], center_2[2], col = 3, pch = 4)
```

####(f) In your plot from (a), color the observations according to the
####cluster labels obtained.

```{r f, comment=FALSE, warning=FALSE}
plot(clusterData[, 1], clusterData[, 2], col=(labels + 1), pch = 20, ceclusterData = 2)
```

###III
####(a) Inspect the data. Explain the meaning of the item.

```{r 3a, comment=FALSE, warning=FALSE}
library(arules)
data("AdultUCI") 
dim(AdultUCI) 
AdultUCI[1:2,]
AdultUCI[["fnlwgt"]] <- NULL 
AdultUCI[["education-num"]] <- NULL 
AdultUCI[[ "age"]] <- ordered(cut(AdultUCI[[ "age"]], c(15,25,45,65,100)), labels = c("Young", "Middle-aged", "Senior", "Old"))  
AdultUCI[[ "hours-per-week"]] <- ordered(cut(AdultUCI[[ "hours-per-week"]], c(0,25,40,60,168)), labels = c("Part-time", "Full-time", "Over-time", "Workaholic")) 
AdultUCI[[ "capital-gain"]] <- ordered(cut(AdultUCI[[ "capital-gain"]], c(-Inf,0,median(AdultUCI[[ "capital-gain"]][AdultUCI[[ "capital-gain"]]>0]), Inf)), labels = c("None", "Low", "High"))  
AdultUCI[[ "capital-loss"]] <- ordered(cut(AdultUCI[[ "capital-loss"]], c(-Inf,0, median(AdultUCI[[ "capital-loss"]][AdultUCI[[ "capital-loss"]]>0]), Inf)), labels = c("None", "Low", "High"))
Adult <- as(AdultUCI, "transactions") 
Adult
```

####(b) Plot the frequencies of items in the data set with support greater than 10%. 

```{r 3b, comment=FALSE, warning=FALSE}
itemFrequencyPlot(Adult, support = 0.1, cex.names = 0.8)
```

####(c) How many rules are there with a minimum support of 1% and a confidence of 60%? 

```{r 3c, comment=FALSE, warning=FALSE}
rules <- apriori(Adult, parameter = list(support = .01, confidence = 0.6, target ="rules"))
```
##### There are 276443 rules.


####(d) Find all the rules related to the variable income. 
####That is, find the factors that are associated with high income, and the factors associated with low income. 

```{r 3d, comment=FALSE, warning=FALSE}
rulesLowInc <- subset(rules, rhs %in% "income=small" )
rulesHighInc <- subset(rules, rhs %in% "income=large")
```

####(e) If there are redundant rules, prune them.

```{r 3e, comment=FALSE, warning=FALSE}
#prune rules for support = .1 for manageable size
rules <- apriori(Adult, parameter = list(support = .1, confidence = 0.6, target ="rules"))
matrix_sub <- is.subset(rules, rules)
matrix_sub[lower.tri(matrix_sub, diag=T)] <- NA
redundantRules <- colSums(matrix_sub, na.rm=T) >= 1
rules_cut <- rules[!redundantRules]
```

####(f) Inspect the three item sets with highest lift that indicate the low income and high income respectively. 

```{r 3f, comment=FALSE, warning=FALSE}
inspect(sort(rulesLowInc, by = "lift")[1:3])
inspect(sort(rulesHighInc, by = "lift")[1:3])
```

####(g) Plot your rules with support and confidence.

```{r 3g, comment=FALSE, warning=FALSE}
library(arulesViz)
plot(rules)
```